Frequentist Modelling

### 2024-09-18 notebooks/02-frequentist-modelling.qmd
Bayesian modelling done and ready to integrate to the manuscript

---
title: Modelling results
subtitle: "Are there unconscious visual images in aphantasia? Development of an implicit priming paradigm"
---

@@

group_effects <- 
  df_questionnaires |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia, data = data)),
    lm_ranked = list(lm(rank(score) ~ aphantasia, data = data)),
    ranked_means = list(prepare_means(lm_ranked, questionnaire)),
    table = list(generate_model_summary(lm, lm_ranked, questionnaire))
  )

@@

  # Long format first -----------------------------------------------------
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # Plotting --------------------------------------------------------------
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = aphantasia, 
    color = aphantasia
    )) +

@@

  # Scores ---------------------------------------------------------------
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = aphantasia
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # Stars and lines --------------------------------------------------------
  # stars for significance
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 1, y = .95, size  = 10 
    ) +
  # dotted line that separates the different questionnaires
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  # stars 
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 2, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 3, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 4, y = .95, size  = 10 
    ) +

  # Aesthetics -------------------------------------------------------------
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Aphantasic")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Aphantasic")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial")
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8)
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

# ggsave(here("figures/questionnaires.png"), dpi = 600)
```
:::

@@

We will quickly check for accuracy differences between groups with a logistic regression model.

@@

df_questionnaires |> 
  group_by(aphantasia, sub_group) |> 
  count() |> 
  arrange(sub_group) |> 
  rename(Aphantasia = 1, "Sub-group" = 2, N = 3) |> 
  display()
```
:::

@@

@@

# Anova(fit_finer_i, type = 3)
```
:::

@@

# Anova(fit_finer_e, type = 3)
```
:::

>>>>

---
title: Frequentist Modelling
subtitle: "Are there unconscious visual images in aphantasia? Development of an implicit priming paradigm"
---

@@

group_effects <- 
  df_questionnaires |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia + age, data = data)),
    lm_ranked = list(lm(rank(score) ~ aphantasia + age, data = data)),
    ranked_means = list(prepare_means(lm_ranked, questionnaire)),
    table = list(generate_model_summary(lm, lm_ranked, questionnaire))
  )

@@

  # -------------------------- Long format first
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # -------------------------- Plotting
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = aphantasia, 
    color = aphantasia
    )) +

@@

  # --------------------- Scores
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = aphantasia
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # --------------------- Stars and lines
  # stars for significance
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 1, y = .95, size  = 10 
    ) +
  # dotted line that separates the different questionnaires
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  # stars 
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 2, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 3, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 4, y = .95, size  = 10 
    ) +

  # --------------------- Aesthetics
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Aphantasia")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Aphantasia")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial")
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8)
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

ggsave(here("figures/questionnaires.png"), dpi = 600)
```
:::

@@

We quickly checked for accuracy differences between groups with a logistic regression model.

@@

df_i_acc |> 
  select(subjectid, aphantasia, sub_group) |> 
  distinct() |> 
  group_by(aphantasia, sub_group) |> 
  count() |> 
  arrange(sub_group) |> 
  rename(Aphantasia = 1, "Sub-group" = 2, N = 3) |> 
  display()
```
:::

@@

## Questionnaires analyses

::: {#supptbl-questionnaire-results}
```{r}
#| label: table-finer-questionnaires
#| echo: false

group_effects_finer <- 
  df_questionnaires |>
  filter(sub_group != "Hyperphantasia") |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ sub_group + age, data = data)),
    lm_ranked = list(lm(rank(score) ~ sub_group + age, data = data)),
    means = list(generate_model_means(lm, questionnaire, group = "sub_group")),
    contrasts = list(
      estimate_contrasts(lm, contrast = "sub_group", p_adjust = "none") |> 
        mutate(Questionnaire = questionnaire)
        ),
    # table = list(generate_model_summary(lm, lm_ranked, questionnaire, group = "sub_group"))
  )

bind_rows(group_effects_finer$means) |> display()
bind_rows(group_effects_finer$contrasts) |> display()
```
:::

```{r}
#| label: questionnaire-plot-setup-finer
#| echo: false

norm_means_finer <- 
  df_q_norm |> 
  filter(sub_group != "Hyperphantasia") |>
  group_by(sub_group) |>
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "score"
  ) |> 
  group_by(Questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ sub_group, data = data)),
    means = list(estimate_means(lm, by = "sub_group"))
  ) |> 
  unnest(means) |> 
  select(!c(data, lm)) |> 
  ungroup() |> 
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire,
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    ))

norm_means_finer
```

::: {#suppfig-questionnaire-plot}

```{r}
#| label: plot-questionnaires-finer
#| fig-width: 10
#| fig-height: 5

df_q_norm |> 
  filter(sub_group != "Hyperphantasia") |>

  # --------------------- Long format first
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # --------------------- Plotting 
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = sub_group, 
    color = sub_group
    )) +

  # ---------------- Scores
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means_finer,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = sub_group
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # ---------------- Lines
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +

  # --------------- Significance
  # VVIQ control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 0.83, y_star = .93,
    x_line = 0.66, x_line_end = 1,
    y_line = .92
  ) +
  geom_segment(
    x = 0.66, xend = 1.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +
  # VVIQ hypo vs aph
  plot_significance_labels(label = "***", size = 7,
    x_star = 1.17, y_star = .39,
    x_line = 1, x_line_end = 1.34,
    y_line = .38
  ) +

  # SUIS control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 1.83, y_star = .93,
    x_line = 1.66, x_line_end = 2,
    y_line = .92
  ) +
  geom_segment(
    x = 1.66, xend = 2.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +
  # SUIS hypo vs aph
  plot_significance_labels(label = "*", size = 7,
    x_star = 2.17, y_star = .39,
    x_line = 2, x_line_end = 2.34,
    y_line = .38
  ) +

  # OSIQ-O control vs others
  plot_significance_labels(label = "***", size = 7,
    x_star = 2.83, y_star = .93,
    x_line = 2.66, x_line_end = 3,
    y_line = .92
  ) +
  geom_segment(
    x = 2.66, xend = 3.34, 
    y = .91, yend = .91, linewidth = .5, color = "black") +

  # OSIQ-S control vs aph only
  plot_significance_labels(label = "***", size = 7,
    x_star = 4, y_star = .84,
    x_line = 3.66, x_line_end = 4.34,
    y_line = .83
  ) +
  # OSIQ-S hypo vs aph
  plot_significance_labels(label = "*", size = 7,
    x_star = 4.17, y_star = .78,
    x_line = 4, x_line_end = 4.34,
    y_line = .77
  ) +

  # ----------------------------- Aesthetics 
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Hypophantasia   ", " Aphantasia")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Hypophantasia   ", " Aphantasia")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial"),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8),
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

ggsave(here("figures/questionnaires-finer.png"), dpi = 600)
```
:::

## Response Times analysis

@@

Anova(fit_finer_i, type = 3)
```
:::

@@

Anova(fit_finer_e, type = 3)
```
:::
